---
title: "COVID-19 cases and deaths"
date: "12/3/2021"
output: github_document 
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(broom)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Tracking the course of the COVID pandemic
We will use NY Times COVID tracking data from the very beginning of COVID onset to refresh what has happened and what has changed.
```{r}
us_states <- 
  us_states_long %>%
  filter(date > as.Date("2020-03-01")) %>%
  pivot_wider(names_from = "data_type", 
              values_from = "value") %>%
  rename(state = location) %>%
  select(date, state, cases_total, deaths_total) %>%
  mutate(state = as_factor(state)) %>%
  arrange(state, date) %>%
  group_by(state) %>%
  mutate(
    cases_7day = (cases_total - lag(cases_total, 7)) / 7,
    deaths_7day = (deaths_total - lag(deaths_total, 7)) / 7)

us <- us_states %>%
  group_by(date) %>%
  summarize(across(
    .cols = where(is.double),
    .fns = function(x) sum(x, na.rm = T),
    .names = "{col}"
  ))

us[10:20, ] %>%
  knitr::kable()

# Hereby, we aggregate the cases and deaths in the unit of 7-day and also added up for national levels. Now we are wondering what is the relationship between positive cases and death? How has it varied as the pandemic has progressed? 
```
Through a line graph showing trends in cases and deaths in the U.S, we are able to tentatively examine and hypothesize the correlation between the weekly cases and deaths. 
```{r}
coeff <- 20   
# Since deaths_7day is too small to plot at the same time with cases, we add a random number to make deaths_7day more visible, "10" is not work so well, ~30 seems more suitable 

us %>%
  ggplot(aes(date, cases_7day)) +
  geom_line(color = "orange") +
  theme(legend.position = "none") +
  geom_line(aes(x = date, y = deaths_7day * coeff), color = "red") +
  scale_y_continuous(
    labels = scales::comma,
    name = "Cases",
    sec.axis = sec_axis(deaths_7day ~ . / coeff,
      name = "Deaths",
      labels = scales::comma
    )
  ) +
  theme(
    axis.title.y = element_text(color = "orange", size = 13),
    axis.title.y.right = element_text(color = "red", size = 13)
  ) +
  labs(
    title = "U.S. Cases vs. Deaths",
    subtitle = "7-Day Average",
    caption = "Source: NY Times, Arthur Steinmetz",
    x = "Date"
  )
```


## 2. Trends in cases and deaths
While the line graph shows that the more cases, the less deaths, it is not a reasonable relationship to conclude. So we turn to use a simple regression to further our question. 
```{r}
lm_df = 
  lm(deaths_7day ~ cases_7day + date, data = us) %>%
  broom::tidy() 

# While the predictors both have significant p-value"date" weights more than "cases_7day" in predicting the deaths_7day. Therefore, we suggest to adjust the deaths versus cases by time series if the deaths want to be predicted by cases weekly.
```
Overall, the death rate has lowered as well as the positive cases has slowed down/ turned from its exponential increment. But how does this happen? Can we ascribe this to the accessibility of vaccination?

## 3. The effectiveness of vaccine
By taking calculated incidence rates from CDC, which suggests cases among persons who were not fully vaccinated compared with those among fully vaccinated persons decreased from 11.1 to 4.6, we try to model a hypothesized comparison between cases trend line from the onset of pandemic and see how that would bend when the very first vaccine is available. 

CDC Reference: Scobie HM, Johnson AG, Suthar AB, et al. Monitoring Incidence of COVID-19 Cases, Hospitalizations, and Deaths, by Vaccination Status — 13 U.S. Jurisdictions, April 4–July 17, 2021. MMWR Morb Mortal Wkly Rep 2021;70:1284–1290. DOI: http://dx.doi.org/10.15585/mmwr.mm7037e1external icon.
```{r vaccine before}
# assuming 11.1% increase weekly
before_vac <- 0:284/7
#from 2020 March 1st to 12.14 the first vaccine available in the US=9*30 + 14

  cases =
    round(104*(1.11^before_vac), digits = 0) 
#104 cases from above US aggregation data on 2020 3.2
  
  rbind(before_vac, cases)

# and the number of increasing 
  ggplot() + 
    geom_point(aes(x = before_vac, y = cases)) + 
    geom_line(aes(x = before_vac, y = cases)) + 
    xlab("Day") + 
    ylab("Total Cases")
```

```{r after vaccine}
#assuming 4.6% increase weekly after vaccine for another 100 days 
after_vac <- 285:385/7

# data for without vaccine
without_vac <- c(before_vac, after_vac)
cases_after <- round(104*(1.11^without_vac), digits = 0)

# data with vaccine
case_vac.285to300 =
    round(cases[285]*(1.046^(1:101)), digits = 0) 

# plot comparing hypothesized increase for before and after vaccine
  ggplot() + 
    geom_line(aes(x = without_vac,
                  y = cases_after)) +
    geom_point(aes(x = before_vac, 
                   y = cases)) + 
    geom_line(aes(x = after_vac, 
                  y = case_vac.285to300), col = "blue") +
    xlab("Day") + 
    ylab("Total Cases")

  # Apparently without any public health control, cases are hard to slow down but does our hypothesized bending curve with vaccine available corresponding to the actual cases number? If it is very different, how long does vaccine take its effectiveness? How does the distribution of vaccine by sex, age, and race etc. demographic characteristics impact on the effectiveness of vaccine distribution and containing cases increment? 
```

```{r}
case_vac.285to300 
```

